### 2023-03-15 outputs/paper/paper.qmd
skeleton for the quarto doc

---
title: "Women work education happiness lol"
subtitle: "insert here lol"
author: 
  - Jenny Shen
  - Joyce Xuan
  - Jason Ngo
thanks: "Code and data are available at: INSERT GITHUB."
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
format: 
  pdf:
    documentclass: article
    geometry: margin = 1in
    abstract: "hello"
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

#### Initialize quarto workspace by reading csv's ####
library(tidyverse)
library(here)

gss2000_cleaned <- read.csv(here::here("outputs/data/gss2000_cleaned.csv"))
gss2002_cleaned <- read.csv(here::here("outputs/data/gss2002_cleaned.csv"))
gss2004_cleaned <- read.csv(here::here("outputs/data/gss2004_cleaned.csv"))
gss2006_cleaned <- read.csv(here::here("outputs/data/gss2006_cleaned.csv"))
gss2008_cleaned <- read.csv(here::here("outputs/data/gss2008_cleaned.csv"))
gss2010_cleaned <- read.csv(here::here("outputs/data/gss2010_cleaned.csv"))
gss2012_cleaned <- read.csv(here::here("outputs/data/gss2012_cleaned.csv"))
gss2014_cleaned <- read.csv(here::here("outputs/data/gss2014_cleaned.csv"))
gss2016_cleaned <- read.csv(here::here("outputs/data/gss2016_cleaned.csv"))
gss2018_cleaned <- read.csv(here::here("outputs/data/gss2018_cleaned.csv"))
gss2021_cleaned <- read.csv(here::here("outputs/data/gss2021_cleaned.csv"))
merged_gss <- read.csv(here::here("outputs/data/merged_gss.csv"))

```

# Introduction
text goes here

# Data

## Data Source and Methodology
text goes here

## Attributes
text goes here

## Analysis
text goes here

### Mean Education Level from 2000 - 2021
```{r}
#| label: fig-1
#| fig-cap: caption goes here
#| echo: false
#| warning: false

library(tidyverse)

## mean education level over 20 yrs ##

merged_gss |> group_by(Year) |> summarize(mean= mean(Highest_Educ_Level, na.rm = TRUE)) |>
  ggplot(aes(x = Year, y = mean)) +
  geom_line(group = 1) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
  labs(
    title = "Mean highest education level of Women over 20 Years",
    x = "Year",
    y = "Mean Educ Level %"
  ) 
```
text goes here

### Mean Occupational Prestige from 2000 - 2021
```{r}
#| label: fig-2
#| fig-cap: caption goes here
#| echo: false
#| warning: false

## mean occ. prestige pct over 20 years ##

merged_gss |> group_by(Year) |> summarize(mean = mean(Occupational_Prestige_Pct, na.rm = TRUE)) |>
  ggplot(aes(x = Year, y = mean)) +
  geom_line(group = 1) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
  labs(
    title = "Occupational Prestige of Women over 20 Years",
    x = "Year",
    y = "Mean Occ. Prest. %"
  ) 
```
text goes here

### Women's Happiness from 2000 - 2021
```{r}
#| label: fig-1
#| fig-cap: caption goes here
#| echo: false
#| warning: false

merged_gss |> group_by(Year) |> count(Happiness) |>
  ggplot(aes(x = Year, y = n, fill = Happiness)) +
  geom_col(position = "dodge") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
  labs(
    title = "Womens Happiness over 20 Years",
    x = "Year",
    y = "Count"
  )
```
text goes here

# Results
sum up all 3 graphs here

# Discussion

## Womens Education Discussion
jenny section

## Womens Occupational Prestige Discussion
joyce section

## Womens Happiness Discussion
jason section

## Biases and Ethical Concerns
jason

## Weaknesses and next steps
joyce

\newpage

\appendix

# Appendix {-}

our survey is here lol

\newpage

# References